Synchrotron X-Ray Rims in Tycho's Supernova Remnant are Energy Dependent

### 2014-09-20 manuscript/paper-tycho.tex
Paper table code; add/fix up paper tables
Working code to generate/format paper tables in ipynb.  Add tables for
individual regions w/ reduced chi-sqrs, tables w/ averaged fit params for all
filaments.

% This paper
\newcommand*{\tsup}{\textsuperscript}
\newcommand*{\Chandra}{\textit{Chandra}\ }
\newcommand*\mean[1]{\bar{#1}}
\newcommand*{\tsynch}{\tau_{\mt{synch}}}
\defcitealias{ressler2014}{R14}

@@

@@

Explain how the transport models constrain diffusion via $m_E$, define $m_E$.
Explain how we back out a filament shape/width from these models

@@

@@

We selected 13 regions in 5 distinct filaments around Tycho's shock
(Figure~\ref{fig:snr}) based on the following criteria:
(1) filaments should be clear of spatial plumes of thermal ejecta in \Chandra
images; this rules out, e.g., areas of strong thermal emission on Tycho's eastern limb.
(2) filaments should be singular and localized; that is, multiple filaments
should either not overlap or completely overlap (rules out parts of NE limb);
(3) filaments should have clear FWHMs bands; i.e., a peak should be evident
above the background signal or downstream thermal emission (rules out some
faint southern wisps) \note[Aaron]{wishy-washy qualitative}.

% NOTE generate final, higher quality figure + newer figure for regions-5
\begin{figure}
    \centering
    %\plotone{figures/f0-snr.pdf}
    \includegraphics[width=0.35\textwidth]{figures/f0-snr.pdf}
    \caption{RGB image of Tycho with region selections overlaid.  Bands are
    0.7--1 keV (red), 1--2 keV (green) and 2--7 keV (blue).
    \note[Aaron]{Could show the spectrum extraction regions on here too?
    Maybe too cluttered.}}
    \label{fig:snr}
\end{figure}

@@

We fit $0.5$--$7 \unit{keV}$ spectra from each region's upstream and downstream
sections to an absorbed power law model (XSPEC 12.8.1, \texttt{phabs*po}) with
spectral index $n$, hydrogen column density $n_H$, and a normalization as free
parameters.  Table~\ref{tab:spec} lists best fit parameters and reduced
$\chi^2$ values for all regions.  Spectra from thin rims (upstream sections)
are well-fit by the power law model alone; spectra from sections downstream of
the rims are generally poorly-fit, reflecting thermal contamination primarily
from \ion{Si}{13} and \ion{S}{15} He$\alpha$ line emission at $1.85$ and $2.45$
keV.  Although most downstream spectra show thermal contamination, some regions
(e.g., 4, 5 on Tycho's northwestern limb) are reasonably well-fit downstream by
an absorbed power law as well due to a larger distance between the forward
shock and contact discontinuity.
Upstream spectra fits spectral indices between $2.5$--$3$ are consistent with
synchrotron radiation spectra for ???? population of electrons
\note[Aaron]{citation for e- distr. indices?}. Best fit column densities from
upstream spectra, between $0.61$--$0.78 \times 10^{22} \unit{cm^{-2}}$, are
consistent with previous fits to Tycho's nonthermal rims \citep{hwang2002} but
somewhat larger than $\abt0.35$--$0.5 \times 10^{22}$ from radio and optical??
HI absorption line measurements \citep{black1984, albinson1986, kothes2004}.
% NOTE keep stated ranges of spectral indices, column densities up to date

\note[Aaron]{On rob's suggestion of fitting thermal model -- can we get away
with just fitting line emission and constraining from there? or is the
whole NEI apparatus needed.  If we add thermal components -- report spectral
indices and reconfirm spectral softening?}

% NOTE keep region numbers updated (in discussion of good/bad downstream spectra)
\begin{table}
    \scriptsize
    \centering
    \caption{Absorbed power law fit parameters for all regions' upstream and
    downstream spectra.  The reduced $\chi^2$ values for upstream spectrum fits
    are consistent with a lack of thermal emission, but values for downstream
    fits indicate a poor fit due to line emission.
    \label{tab:spec}}
    \input{tab-spec.tex}
\end{table}

@@

In Table~\ref{tab:fwhms} we report FWHM measurements for all of our regions.
As previously noted, we could not resolve FWHMs for some regions in the 0.7--1 keV band.
NOTE very low counts/numbers of regions for some averages (typically 1--2, see
Figure~\ref{fig:snr}).

Mention the number of counts in individual profiles?
May affect whether we go back and take smaller regions, or not.

% TODO give individual regions' FWHMs and m_E values
\begin{table*}[ht]
    \tiny
    \centering
    \caption{FWHMs for all regions.\label{tab:fwhms}}
    %\input{tab-fwhms-ave-arith.tex}
    %\input{tab-fwhms-ave-geom.tex}
    \input{tab-fwhms.tex}
\end{table*}

% -------------------------
% Model fit results, tables
% -------------------------
\subsection{Model fit results}
\label{sec:fit-results}
% TODO do all the work, but with different FWHM calculation -- to confirm
%      that it doesn't make a big difference...

% TODO keep updated -- give tables for ONE region with both simple, full fits
We find that the simpler catastrophic dump model (equation~\eqref{eq:simp-mod})
produces the same qualitative trends seen in the full continuous energy loss
model.
The first row of Table~\ref{tab:fits} compares simple and full model fits for
Region 1(?).
The key differences are \emph{blah}.  The errors are comparable, the parameters
are comparable but shifted by X amount; this is consistent in all filaments?
(details tbd).  Consequently, we present only the full model fits for all other
filaments in Table~\ref{tab:fits}.  However, the simple model is useful for
building intuition and we will refer to its results in the following discussion
(?); \citetalias{ressler2014} also give examples of simple model fits in SN
1006 \note[Aaron]{sentence maybe unnecessary}.

% Individual region fit results
Table~\ref{tab:fits} presents full model fit results for all regions
considered. Errors are computed by varying free parameters $B_0$,
$\eta_2$ to obtain $\Delta\chi^2 = 1$, corresponding to $1$-$\sigma$ confidence
limits on our fit parameters.

@@

% TODO Assemble table of regions
\begin{table*}[ht]
    \scriptsize
    \centering
    \caption{Full model best fits for individual regions, Filament 1.
    \label{tab:fits}}
    \input{tab-fits-flmt1.tex}  % TODO CHANGE/UPDATE THIS
\end{table*}

In Figure~\ref{fig:fits} we plot the full model fits for the same regions
tabulated in Table~\ref{tab:fits}

% TODO Assemble all the individual plots.
\begin{figure}[ht]
    \centering
    \plottwo{figures/f0-fits-avg-flmt1.png}{figures/f0-fits-avg-flmt2.png} \\
    \plottwo{figures/f0-fits-avg-flmt3.png}{figures/f0-fits-avg-flmt4.png} \\
    \plottwo{figures/f0-fits-avg-flmt5.png}{figures/f0-box.png}
    \plottwo{figures/f0-fits-avg-flmt1-geom.png}{figures/f0-fits-avg-flmt2-geom.png} \\
    \plottwo{figures/f0-fits-avg-flmt3-geom.png}{figures/f0-fits-avg-flmt4-geom.png} \\
    \plottwo{figures/f0-fits-avg-flmt5-geom.png}{figures/f0-box.png}
    \caption{Plots of best-fits to averaged FWHMs (arithmetic, geometric avg).
    \note[Aaron]{Need better subplot display in final version, if we keep these}}
    \label{fig:fits-all}
\end{figure}

@@

\begin{table*}
    \scriptsize
    \centering
    \caption{Best fits with $\eta_2 = 1$ fixed, for filament averaged FWHMs.
    \label{tab:fits-avg-eta2-fix}}
    \input{tab-fits-avg-eta2-fix.tex}
\end{table*}

@@

\begin{table*}
    \scriptsize
    \centering
    \caption{Filament-wide average of best-fit parameters for constituent regions.
    \label{tab:fits-avg}}
    \input{tab-fits-par-avg.tex}
\end{table*}

@@

It would be helpful to average $m_E$ values together, under the
assumption that they should be somewhat consistent within a given filament.
Then report the values and errors observed.

Recall that our FWHMs (Table~\ref{tab:fwhms}) have a large range of values,
IF magnetic damping is relevant in controlling filament widths, we might, e.g.,
expect a maximum filament width at which $m_E$ sharply levels off, being now
restricted by the magnetic field turn off.  We simply do not observe this in
our data between $0.7$--$7 \unit{keV}$.  We must conclude that, as in SN 1006
\citepalias{ressler2014}, magnetic damping cannot be relevant to our
observations.

@@

For comparison to \citetalias{ressler2014}, we reproduce Sean's table
(Table~\ref{tab:sean} and present full model fits with 2 and 3 energy bands
(Tables~\ref{tab:sn1006-2band} and \ref{tab:sn1006-3band}, respectively).
We don't compare simple model results, as they are identical (only the error
calculations differ).
The procedure of Tables~\ref{tab:sean},~\ref{tab:sn1006-2band} are not quite
the same.  Sean used $m_E$ at 2 keV instead of the $1$--$2$ keV width for
fitting, whereas I used both widths.  But, it should not matter much as
the fit often has $\chi^2 \sim 0$, either way (depending on the exact
measurements and values of $\mu$).

@@

% ------------------
% Sean's Cas A table
% ------------------
\section{Sean's Cas A comparison (DRAFT ONLY)}

This is dredged from the \LaTeX source for Sean's paper, available on arXiv.
I put this here just for my own reference.

A detailed application of our results to other SNRs such as Cas A will
require much more extensive analysis, but we can use the published
filament widths of Araya et al. (2010) for Cas A to get preliminary
estimates of the magnetic field strength and diffusion coefficient by
applying our model. In their data, it appears that the filaments in
Cas A shrink by a factor of $\sim 0.8$ between 0.3 and 3 keV, while the
filament widths appear to be energy-independent between 3 and 6 keV.
Qualitatively, this is consistent with the loss-limited model, as our
parameter $m_{\rm E}$ is predicted to decrease with energy.  For the
lower energy range of 0.3--3 keV, reproducing $m_{\rm E}\sim -0.1$
(equivalent to the factor of 0.8 drop in size) requires magnetic
fields on the order of 200-500 $\mu$G and diffusion coefficients about
$5 \times D_{\rm Bohm}$(3 keV), about an order of magnitude higher
than the values one obtains by neglecting the energy dependence. One
can also see directly from Figure~\textbf{?!} that $\mu < 1$ models
of the diffusion coefficient are excluded for $m_{\rm E}\sim -0.1$.

To quickly compare how our results will differ from those of other authors, we
applied the analytic approximation of Equation 22 and Equation 9 to the FWHMs
presented by Araya et al. (2010) for Cas A. Here we used $m_{\rm E} = \log(
{\rm FWHM}(3keV)/{\rm FWHM}(.3keV))/\log(10)$ and fit the filaments at a photon
energy of 3 keV.  Hence $\eta_3 \equiv D/D_{\rm Bohm}(3 keV)$. Using the same
parameters, our results are shown in Table~\ref{tab:araya}. To fit the energy
dependence of the FWHM (i.e. $m_{\rm E}$), we required higher diffusion
coefficients and, consequently, magnetic fields about an order of magnitude
higher than their results. We note here that values of $\mu<1$ were unable to
reproduce the data and that the diffusion coefficients cited by Araya et al.
(2010) are below the minimally allowed Bohm value ($\eta_3 = 1$).

We intended this comparison solely as a qualitative analysis, and thus did not
rigorously keep track of uncertainties or extend the comparison to the full
numerical calculation. The conclusion is clear, however, which is that
considering the measured value of $m_{\rm E}$ can have a dramatic effect.

\begin{table}[h]
\tiny
\centering
\caption{Best fit parameters for the filaments of Cas A based on data from
Araya et. al (2010) in varying values of $\mu$ (Approximate Analytic Results)
Dashes denote places where fits were unobtainable.
\label{tab:araya}}
\input{tab-sean-casA.tex}
\end{table}

>>>>

% This paper
\newcommand*{\tsup}{\textsuperscript}
\newcommand*{\Chandra}{\textit{Chandra}\ }
\newcommand*\mean[1]{\bar{#1}}
\newcommand*{\tsynch}{\tau_{\mt{synch}}}
\newcommand*{\mE}{m_\mathrm{E}}
\defcitealias{ressler2014}{R14}

@@

% TODO GENERAL WARNING -- DATA ARE NOT ALL FULLY CONSISTENT AT THIS TIME.
% (spectra/profiles/fwhms/fits).  Will be corrected / good to go in final
% revision.

@@

Explain how the transport models constrain diffusion via $\mE$, define $\mE$.
Explain how we back out a filament shape/width from these models

@@

Discuss $\mE$, I have NOT defined this quantity below.  % TODO

@@

We selected 13 regions in 5 distinct filaments around Tycho's shock
(Figure~\ref{fig:snr}) based on the following criteria:
(1) filaments should be clear of spatial plumes of thermal ejecta in \Chandra
images; this rules out, e.g., areas of strong thermal emission on Tycho's eastern limb.
(2) filaments should be singular and localized; that is, multiple filaments
should either not overlap or completely overlap (rules out parts of NE limb);
(3) filaments should have clear FWHMs bands; i.e., a peak should be evident
above the background signal or downstream thermal emission (rules out some
faint southern wisps) \note[Aaron]{wishy-washy qualitative}.
We grouped regions into filaments by visual inspection of the remnant.

% NOTE generate final, higher quality figure + newer figure for regions-5
\begin{figure}
    \centering
    \plotone{figures/f0-snr-inv.png}
    \caption{RGB image of Tycho with region selections overlaid.  Bands are
    0.7--1 keV (red), 1--2 keV (green) and 2--7 keV (blue).
    \note[Aaron]{Could show the spectrum extraction regions on here too?
    Maybe too cluttered.  Figure colors inverted to save ink.}}
    \label{fig:snr}
\end{figure}

@@

We fit $0.5$--$7 \unit{keV}$ spectra from each region's upstream and downstream
sections to an absorbed power law model (XSPEC 12.8.1, \texttt{phabs*po}) with
spectral index $n$, hydrogen column density $n_H$, and a normalization as free
parameters.  Table~\ref{tab:spec} lists best fit parameters and reduced
$\chi^2$ values for all regions.  Spectra from thin rims (upstream sections)
are well-fit by the power law model alone; spectra from sections downstream of
the rims are generally poorly-fit, reflecting thermal contamination primarily
from \ion{Si}{13} and \ion{S}{15} He$\alpha$ line emission at $1.85$ and $2.45$
keV.  Although most downstream spectra show thermal contamination, some regions
(e.g., 4, 5 on Tycho's northwestern limb) are reasonably well-fit downstream by
an absorbed power law as well due to a larger distance between the forward
shock and contact discontinuity.
Upstream spectra fits spectral indices between $2.5$--$3$ are consistent with
synchrotron radiation spectra for ???? population of electrons
\note[Aaron]{citation for e- distr. indices?}. Best fit column densities from
upstream spectra between $0.61$--$0.78 \times 10^{22} \unit{cm^{-2}}$ are
consistent with previous fits to Tycho's nonthermal rims \citep{hwang2002} but
somewhat larger than $\abt0.35$--$0.5 \times 10^{22} \unit{cm^{-2}}$ from radio
and optical??  HI absorption line measurements \citep{black1984, albinson1986,
kothes2004}.
% NOTE keep stated ranges of spectral indices, column densities up to date

\note[Aaron]{On Rob's suggestion of fitting thermal model -- can we get away
with just fitting line emission and constraining from there? or is the
whole NEI apparatus needed.  If we add thermal components -- report spectral
indices and reconfirm spectral softening?}

% NOTE keep region numbers updated (in discussion of good/bad downstream spectra)
\begin{table}
    \scriptsize
    \centering
    \caption{Region spectra fit parameters\label{tab:spec}}
    \input{tab-spec.tex}
    \tablecomments{Absorbed power law fit parameters are photon index $\Gamma$
    and hydrogen column density $\mathrm{N_H}$.}
\end{table}

@@

In Table~\ref{tab:fwhms} we report FWHM measurements for all of our regions.
As previously noted, we could not resolve FWHMs for some regions in the
0.7--1 keV band.  We also report $\mE$ values for all but the lowest energy
band, computed point-to-point as $\mE(E_2) = \ln(w_2/w_1) / \ln(E_2/E_1)$ where
$w_1, w_2$ and $E_1, E_2$ are FWHMs and lower energy values for each energy
band (e.g., $\mE(\text{1 keV}) = \ln(w(\text{1-1.7 keV}/w(\text{0.7-1 keV})) /
\ln(\text{1 keV}/\text{0.7 keV})$ is computed from widths/energies.  Although
the values of $\mE$ reflect a good deal of variability in the underlying width
measurements, we see that most regions and energy bands display a decrease in
rim widths as a function of energy.

\begin{table*}[ht]
    \tiny
    \centering
    \caption{Measured FWHMs for all regions.\label{tab:fwhms}}
    \input{tab-fwhms.tex}
\end{table*}

Here we have presented $\mE$ to demonstrate the fall-off of rim width energy.
However, we emphasize that $\mE$ is currently not factored into our subsequent
analysis.

% -------------------------
% Model fit results, tables
% -------------------------
\subsection{Model fit results}
\label{sec:fit-results}

% TODO do all the work, but with different FWHM calculation -- to confirm
%      that it doesn't make a big difference...

% NOTE keep updated -- for which region are we giving simple and full fits?
We find that the simpler catastrophic dump model (equation~\eqref{eq:simp-mod})
produces the same qualitative trends seen in the full continuous energy loss
model.
The first row of Table~\ref{tab:fits} compares simple and full model fits for
Region 1.
The key differences are \emph{blah}.  The errors are comparable, the parameters
are comparable but shifted by X amount; this is consistent in all filaments?
(details tbd).
Consequently, we present only the full model fits for all other
filaments in Table~\ref{tab:fits}.  However, the simple model is useful for
building intuition and we will refer to its results in the following discussion
(?); \citetalias{ressler2014} also give examples of simple model fits in SN
1006 \note[Aaron]{sentence maybe unnecessary}.

% Individual region fit results
Table~\ref{tab:fits} presents full model fit results for all regions
considered. Errors are computed by varying free parameters $B_0$,
$\eta_2$ to obtain $\Delta\chi^2 = 1$, corresponding to $1$-$\sigma$ confidence
limits on our fit parameters.
\begin{table}[ht]
    \tiny
    \centering
    \caption{Full model best fits for individual regions, Filament 1.
    \label{tab:fits}}
    \input{tab-fits-all.tex}
\end{table}

\begin{table}[ht]
    \tiny
    \centering
    \caption{Full model best fits for individual regions, Filaments 2--5.
    \label{tab:fits-pt2}}
    \input{tab-fits-all-pt2.tex}
\end{table}

@@

In Figure~\ref{fig:fits-all} we plot the full model fits for the same regions
tabulated in Table~\ref{tab:fits}

% NOTE need better subplot display in final version, if we keep this
\begin{figure}[ht]
    \centering
    \plottwo{figures/fits-sep13/reg01.png}{figures/fits-sep13/reg02.png} \\
    \plottwo{figures/fits-sep13/reg03.png}{figures/fits-sep13/reg04.png} \\
    \plottwo{figures/fits-sep13/reg05.png}{figures/fits-sep13/reg06.png} \\
    \plottwo{figures/fits-sep13/reg07.png}{figures/fits-sep13/reg08.png} \\
    \plottwo{figures/fits-sep13/reg09.png}{figures/fits-sep13/reg10.png} \\
    \plottwo{figures/fits-sep13/reg11.png}{figures/fits-sep13/reg12.png} \\
    \plottwo{figures/fits-sep13/reg13.png}{figures/f0-box.png}
    \caption{Fitted full model widths as a function of energy for all regions,
    with measured data; best fit parameters are given in
    Tables~\ref{tab:fits},~\ref{tab:fits-pt2}.  Plots are ordered by region
    number.}
    \label{fig:fits-all}
\end{figure}

@@

\begin{table*}
    \scriptsize
    \centering
    \caption{Best fits with $\eta_2 = 1$ fixed for filament averaged FWHMs.
    \label{tab:fits-avg-eta2-fix}}
    \input{tab-fits-avg-eta2-fix.tex}
\end{table*}

@@

\begin{table*}
    \scriptsize
    \centering
    \caption{Filament averages of best-fit parameters for constituent regions
    \label{tab:fits-avg}}
    \input{tab-fits-par-avg.tex}
\end{table*}

@@

It would be helpful to average $\mE$ values together, under the
assumption that they should be somewhat consistent within a given filament.
Then report the values and errors observed.

Recall that our FWHMs (Table~\ref{tab:fwhms}) have a large range of values,
IF magnetic damping is relevant in controlling filament widths, we might, e.g.,
expect a maximum filament width at which $\mE$ sharply levels off, being now
restricted by the magnetic field turn off.  We simply do not observe this in
our data between $0.7$--$7 \unit{keV}$.  We must conclude that, as in SN 1006
\citepalias{ressler2014}, magnetic damping cannot be relevant to our
observations.

@@

For comparison to \citetalias{ressler2014}, we reproduce Sean's full model fit
table (Table~\ref{tab:sean}) and present full model fits with 2 and 3 energy
bands (Tables~\ref{tab:sn1006-2band} and \ref{tab:sn1006-3band}, respectively).
We don't compare simple model results, as they are identical (only the error
calculations differ).
The procedures of Tables~\ref{tab:sean},~\ref{tab:sn1006-2band} are not quite
the same.  Sean used $\mE$ at 2 keV instead of the $1$--$2$ keV width for
fitting, whereas I used both widths.  But, it should not matter much as
the fit often has $\chi^2 \sim 0$, either way (depending on the exact
measurements and values of $\mu$).

@@
